Unshuffling Permutations

### 2016-02-09 new-article.tex
LATIN 2016 corrections

\newcommand{\LabeledCrossingLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingLL

@@

\newcommand{\LabeledCrossingLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingLR

@@

\newcommand{\LabeledCrossingRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingRL

@@

\newcommand{\LabeledCrossingRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingRR

@@

\newcommand{\LabeledInclusionLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionLL

@@

\newcommand{\LabeledInclusionLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionLR

@@

\newcommand{\LabeledInclusionRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionRL

@@

\newcommand{\LabeledInclusionRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionRR

@@

\newcommand{\LabeledPrecedenceLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceLL

@@

\newcommand{\LabeledPrecedenceLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceLR

@@

\newcommand{\LabeledPrecedenceRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceRL

@@

\newcommand{\LabeledPrecedenceRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceRR

@@

This section is devoted to proving hardness of recognizing square
permutations. In the same way as happens with words, we shall use a
linear graph framework where deciding whether a permutation is a square
reduces to computing some specific matching in the associated linear
graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.

@@

@@

@@

Before proving hardness, we give an easy lemma that will prove extremely
useful for simplifying the proof of upcoming Proposition~\ref{proposition:hardness}.

@@

\begin{lemma}
  \label{lemma:at most one edge monotone}
  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
  and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
  that avoids
  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
  and \CrossingRL, and such that for any two distinct arcs $(a, a')$ and
  $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only
  if $\pi(a') < \pi(b')$.
  The following assertions hold:
  \begin{enumerate}
    \item
    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
    $\sigma_1$ is above $\sigma_2$
    (see Fig.~\ref{subfig:increasing before and above decreasing}),
    then $\vv{\mathcal{M}}$ contains either at most one
    $(\sigma_1, \sigma_2)$-arc or at most one
    $(\sigma_2, \sigma_1)$-arc.
    \item
    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
    is below $\sigma_2$
    (see Fig.~\ref{subfig:decreasing before and below increasing}),
    then $\mathcal{M}$ contains at most one
    $(\sigma_1, \sigma_2)$-arc or at most one $(\sigma_2, \sigma_1)$-arc.
  \end{enumerate}
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
  We only prove the first assertion since the proof for the second one
  is similar.
  Suppose, aiming at a contradiction, that the assertion does not hold.
  Therefore, since $\vv{\mathcal{M}}$ avoids
  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
  and \CrossingRL,
  $\vv{\mathcal{M}}$ contains either two crossing $(\sigma_1, \sigma_2)$-arcs
  or two crossing $(\sigma_2, \sigma_1)$-arcs.
  Suppose first that it contains two crossing $(\sigma_1, \sigma_2)$-arcs
  $(a, a')$ and $(b, b')$ with $a < b$.
  Therefore, we have $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing)
  and $\pi(a') > \pi(b')$ (since
  $\sigma_2$ is decreasing), which contradicts our hypothesis about
  $\vv{\mathcal{M}}$.
  If $\vv{\mathcal{M}}$ contains two crossing $(\sigma_2, \sigma_1)$-arcs
  $(a, a')$ and $(b, b')$ with $a < b$, then
  we have $\pi(a) > \pi(b)$ (since $\sigma_2$ is decreasing)
  and $\pi(a') < \pi(b')$ (since
  $\sigma_2$ is increasing), which contradicts our hypothesis about
  $\vv{\mathcal{M}}$.
  \qed
\end{proof}

>>>>

\newcommand{\LabeledCrossingLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingLL

@@

\newcommand{\LabeledCrossingLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingLR

@@

\newcommand{\LabeledCrossingRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingRL

@@

\newcommand{\LabeledCrossingRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \LabeledCrossingRR

@@

\newcommand{\LabeledInclusionLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionLL

@@

\newcommand{\LabeledInclusionLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionLR

@@

\newcommand{\LabeledInclusionRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionRL

@@

\newcommand{\LabeledInclusionRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
\end{tikzpicture}
}% end \LabeledInclusionRR

@@

\newcommand{\LabeledPrecedenceLL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceLL

@@

\newcommand{\LabeledPrecedenceLR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceLR

@@

\newcommand{\LabeledPrecedenceRL}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceRL

@@

\newcommand{\LabeledPrecedenceRR}[5]{%
\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
                    text height=1.5ex,text depth=.25ex]
\draw node [] (U1)               {#2};
\draw node [] [right of=U1] (U2) {#3};
\draw node [] [right of=U2] (U3) {#4};
\draw node [] [right of=U3] (U4) {#5};
\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
\end{tikzpicture}
}% end \PrecedenceRR

@@

This section is devoted to proving hardness of recognizing square
permutations. In the same way as happens with words, we shall use a
linear graph framework where deciding whether a permutation is a square
reduces to computing some specific matching in the associated linear
graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
We have, however, to deal with oriented perfect matchings.
The needed properties read as follows.

@@

The rationale for introducing properties $\mathbf{P_1}$ and $\mathbf{P_2}$
stems from the following lemma.

@@

% Before proving hardness, we give two easy lemmas that will prove extremely
% useful for simplifying the proof of upcoming Proposition~\ref{proposition:hardness}.

@@

@@

\begin{lemma}
  \label{lemma:at most one edge monotone}
  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
  and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
  that has properties~$\mathbf{P_1}$ and $\mathbf{P_2}$.
  The following assertions hold:
  \begin{enumerate}
    \item
    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
    $\sigma_1$ is above $\sigma_2$
    (see Fig.~\ref{subfig:increasing before and above decreasing}),
    then there is at most one arc between $\sigma_1$ and
    $\sigma_2$ in $\vv{\mathcal{M}}$ (this arc can be a
    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
    \item
    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
    is below $\sigma_2$
    (see Fig.~\ref{subfig:decreasing before and below increasing}),
    then there is at most one arc between $\sigma_1$ and
    $\sigma_2$ in $\vv{\mathcal{M}}$ (this arc can be a
    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
  \end{enumerate}
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
  We only prove the first assertion since the proof for the second one
  is similar.
  Suppose, aiming at a contradiction, that the assertion does not hold.
  Therefore, since $\vv{\mathcal{M}}$ has Property~$\mathbf{P_1}$,
  $\vv{\mathcal{M}}$ contains either two crossing $(\sigma_1, \sigma_2)$-arcs
  or two crossing $(\sigma_2, \sigma_1)$-arcs.
  Suppose first that it contains two crossing $(\sigma_1, \sigma_2)$-arcs
  $(a, a')$ and $(b, b')$ with $a < b$.
  Therefore, we have $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing)
  and $\pi(a') > \pi(b')$ (since
  $\sigma_2$ is decreasing), which contradicts our hypothesis about
  $\vv{\mathcal{M}}$.
  If $\vv{\mathcal{M}}$ contains two crossing $(\sigma_2, \sigma_1)$-arcs
  $(a, a')$ and $(b, b')$ with $a < b$, then
  we have $\pi(a) > \pi(b)$ (since $\sigma_2$ is decreasing)
  and $\pi(a') < \pi(b')$ (since
  $\sigma_2$ is increasing), which contradicts our hypothesis about
  $\vv{\mathcal{M}}$.
  \qed
\end{proof}
    $\pi(a) < \pi(a')$.
        scale=0.3,
      \begin{scope}[]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
      \begin{scope}[yshift=12.5cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
      %
      \begin{scope}[xshift=10cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
    %
      \begin{scope}[xshift=10cm,yshift=12.5cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
  \qquad\;
    $\pi(a) < \pi(a')$.
        scale=0.3,
      \begin{scope}[]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
      \begin{scope}[yshift=12.5cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
      %
      \begin{scope}[xshift=10cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
    %
      \begin{scope}[xshift=10cm,yshift=12.5cm]
        % points
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
        \node [label] (a)  at (1,-1.5)  {$a$};
        \node [label] (ap) at (3,-1.5)  {$a'$};
        \node [label] (b)  at (5,-1.5)  {$b$};
        \node [label] (bp) at (7,-1.5) {$b'$};
        \node (nu1) at (4,-2.5) {$\pi$};
        % arcs
        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
      \end{scope}
  \caption{\label{fig:subfig:value inclusion of arcs}%
    Illustration of Lemma~\ref{lemma:value inclusion of arcs}
    for $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
    and
    $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$.

\begin{lemma}
  \label{lemma:value inclusion of arcs}
  Let $\pi$ be a permutation
  and $\vv{\mathcal{M}}$ be an oriented perfect matching of
  the linear graph $G_\pi$.
  If $\vv{\mathcal{M}}$ has Property $\mathbf{P_2}$ then it avoids patterns
  $$
  \LabeledPrecedenceLL{.3}{$a$}{$a'$}{$b$}{$b'$},
  \qquad
  \LabeledPrecedenceLR{.3}{$a$}{$a'$}{$b$}{$b'$},
  \qquad
  \LabeledPrecedenceRL{.3}{$a$}{$a'$}{$b$}{$b'$},
  \qquad
  \LabeledPrecedenceRR{.3}{$a$}{$a'$}{$b$}{$b'$}
  $$
  for
  (i)
  $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
  and
  $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$
  and
  (ii)
  $\min\{\pi(a), \pi(a')\} > \min\{\pi(b), \pi(b')\}$
  and
  $\max\{\pi(a), \pi(a')\} < \max\{\pi(b), \pi(b')\}$.
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:value inclusion of arcs}]
  We only prove the lemma for case \emph{(i)}
  since the proof for case \emph{(ii)} is similar.
  Therefore, assume
  $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
  and
  $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$,
  and suppose, aiming at a contradiction, that the assertion of the lemma
  does not hold.
  We need to consider two cases.
  For each of the following patterns
  $$
  \LabeledPrecedenceLL{.3}{$a$}{$a'$}{$b$}{$b'$},
  \qquad
  \LabeledPrecedenceLR{.3}{$a$}{$a'$}{$b$}{$b'$},
  $$
  we have
  $\pi(a) < \pi(b)$ and $\pi(a') > \pi(b')$,
  and
  for each of the following patterns
  $$
  \LabeledPrecedenceRL{.3}{$a$}{$a'$}{$b$}{$b'$},
  \qquad
  \LabeledPrecedenceRR{.3}{$a$}{$a'$}{$b$}{$b'$},
  $$
  we have
  $\pi(a) > \pi(b)$ and $\pi(a') < \pi(b')$.
  Hence, it follows that $\vv{\mathcal{M}}$ does not have
  Property $\mathbf{P_2}$. This is a contradiction.
  \qed
\end{proof}

\begin{lemma}
  \label{lemma:for nu_1 and nu_2}
  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
  be a permutation where $\sigma_1$ is an increasing pattern of
  length $k$, $k \geq 3$,  and $\sigma_2$ is below $\sigma_1$,
  and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
  that has properties $\mathbf{P_1}$ and $\mathbf{P_2}$.
  If $\vv{\mathcal{M}}$ contains either a $(\sigma_1, \sigma_2)$-arc
  or a $(\sigma_2, \sigma_1)$-arc,
  then it does not contain a $(\sigma_1, \sigma_1)$-arc.
\end{lemma}
